---
title: "Survival Analysis"
format: html
---

## Libraries
```{r}
library(tidyverse)
library(readxl)
library(survival)
library(survminer)
library(survivalAnalysis)
```


## Load Data
```{r}
cancer_data <- read_excel("cancer_data.xlsx", 
    sheet = "1 Base Data All") |> 
    mutate(
    RACE = ifelse(RACE == "Unknown", "Other", RACE),
    ISDEAD_FLAG = if_else(ISDEAD_FLAG == "y", 0, 1),
    CANCER_GROUP = case_when(
      CANCER_GROUP == "Ca breast" ~ "Breast",
      CANCER_GROUP == "Ca prostate" ~ "Prostate",
      CANCER_GROUP == "Ca female genitalia" ~ "Female Genitalia",
      CANCER_GROUP == "Ca cervix" ~ "Cervix"
      ),
    SURGERY_DATE = `Histopathology Surgical_First_Date`
    ) |> 
  filter(CANCER_GROUP != "Female Genitalia")
```

## Date of first visit to first day of radiotheraphy
```{r}
fir_rad_data <- 
  cancer_data |>
  filter(!is.na(RADIOTHERAPY_FIRST_DATE) &
           !is.na(IALCH_FIRST_VISIT_REG_DATE)) |> 
  mutate(
    first_visit_to_radiotheraphy = abs(round(difftime(RADIOTHERAPY_FIRST_DATE, IALCH_FIRST_VISIT_REG_DATE, units = "weeks"))),
    AGE = floor(time_length(interval(DOB, RADIOTHERAPY_FIRST_DATE), unit = "year")),
    AGE_GROUP = case_when(
      AGE <= 40 ~ "Less than 40",
      TRUE ~ "Greater than 40"
    ),
    EVENT = if_else(DOD < RADIOTHERAPY_FIRST_DATE, 1, 0)
  )
    


surv_object <- Surv(
  time = fir_rad_data$first_visit_to_radiotheraphy,
  event = fir_rad_data$EVENT)

fit.coxph <- coxph(surv_object ~ RACE + GENDER, 
                   data = fir_rad_data)

fir_rad_data |> 
  analyse_multivariate(vars(
    first_visit_to_radiotheraphy,
    ISDEAD_FLAG
  ), vars(GENDER, RACE, AGE_GROUP, CANCER_GROUP
)) |> 
  forest_plot(endpoint_labeller = c(first_visit_to_radiotheraphy = "Date of first visit\nto first day\nof radiotheraphy"))
```

## Ca first visit to chemotheraphy
```{r}
fir_rad_data <- 
  cancer_data |>
    filter(!is.na(CHEMOTHERAPY_FIRST_DATE) & 
             !is.na(CA_FIRST_VISIT_START)) |> 
  mutate(
    first_visit_to_radiotheraphy = abs(round(difftime(CHEMOTHERAPY_FIRST_DATE
, CA_FIRST_VISIT_START
, units = "weeks"))),
    AGE = floor(time_length(interval(
           DOB, CHEMOTHERAPY_FIRST_DATE

         ), unit = "year")),
    AGE_GROUP = case_when(
      AGE <= 40~ "Less than 40",
      TRUE ~ "Greater than 40"
      ),
    EVENT = if_else(DOD < CHEMOTHERAPY_FIRST_DATE, 1, 0)
    )
    


surv_object <- Surv(
  time = fir_rad_data$first_visit_to_radiotheraphy,
  event = fir_rad_data$EVENT)

fit.coxph <- coxph(surv_object ~ RACE + GENDER, 
                   data = fir_rad_data)

fir_rad_data |> 
  analyse_multivariate(vars(
    first_visit_to_radiotheraphy,
    ISDEAD_FLAG
  ), vars(GENDER, RACE, AGE_GROUP, CANCER_GROUP
)) |> 
  forest_plot(endpoint_labeller = c(first_visit_to_radiotheraphy = "Date of cancer first visit\nto first day\nof chemotheraphy"))
```


## Surgery to CA first visit
```{r}
fir_rad_data <- 
  cancer_data |>
  filter(!is.na(CA_FIRST_VISIT_START) &
           !is.na(SURGERY_DATE)) |> 
  mutate(
    first_visit_to_radiotheraphy = abs(round(
      difftime(SURGERY_DATE, CA_FIRST_VISIT_START
, units = "weeks"))),
    AGE = floor(time_length(interval(
           DOB, SURGERY_DATE

         ), unit = "year")),
    AGE_GROUP = case_when(
      AGE <= 40~ "Less than 40",
      TRUE ~ "Greater than 40"
      ),
    EVENT = if_else(DOD < SURGERY_DATE, 1, 0)

    )
    


surv_object <- Surv(
  time = fir_rad_data$first_visit_to_radiotheraphy,
  event = fir_rad_data$EVENT)

fit.coxph <- coxph(surv_object ~ RACE + GENDER, 
                   data = fir_rad_data)

fir_rad_data |> 
  analyse_multivariate(vars(
    first_visit_to_radiotheraphy,
    ISDEAD_FLAG
  ), vars(GENDER, RACE, AGE_GROUP, CANCER_GROUP
)) |> 
  forest_plot(endpoint_labeller = c(first_visit_to_radiotheraphy = "Date of surgery\nto cancer first visit"))
```
## Date of Surgery to Radiotheraphy
```{r}
```{r}
fir_rad_data <- 
  cancer_data |>
  filter(!is.na(RADIOTHERAPY_FIRST_DATE) &
           !is.na(SURGERY_DATE)) |> 
  mutate(
    first_visit_to_radiotheraphy = abs(round(difftime(RADIOTHERAPY_FIRST_DATE, SURGERY_DATE, units = "weeks"))),
    AGE = floor(time_length(interval(DOB, RADIOTHERAPY_FIRST_DATE), unit = "year")),
    AGE_GROUP = case_when(
      AGE <= 40 ~ "Less than 40",
      TRUE ~ "Greater than 40"
    ),
    EVENT = if_else(DOD < RADIOTHERAPY_FIRST_DATE, 1, 0)
  )
    


surv_object <- Surv(
  time = fir_rad_data$first_visit_to_radiotheraphy,
  event = fir_rad_data$EVENT)

fit.coxph <- coxph(surv_object ~ RACE + GENDER, 
                   data = fir_rad_data)

fir_rad_data |> 
  analyse_multivariate(vars(
    first_visit_to_radiotheraphy,
    ISDEAD_FLAG
  ), vars(GENDER, RACE, AGE_GROUP, CANCER_GROUP
)) |> 
  forest_plot(endpoint_labeller = c(first_visit_to_radiotheraphy = "Date of surgery\nto first day\nof radiotheraphy"))
```

